Bayesian Exploration in Deep Reinforcement Learning

### 2022-02-03 papers/main.tex
TEMPORARY DEPLOY COMMIT

@@

\subsection{Noisy Networks}
In NoisyNet \citep{fortunato_noisy_2019}, the Q-function space can be seen
as a stochastic process on the probability space \((\Omega, \mathcal{F}, P)\).
The stochastic process can simply be written as
\(\{Q(\bm{s}, \bm{a}) : (\bm{s}, \bm{a}) \in \state \times \action \}\). For
any \(\omega \in \Omega\), \(Q(\cdot, \cdot, \omega)\) is a sample function
mapping \(\state \times \action \rightarrow \mathbb{R}\). To simplify notation
we will denote sample functions as \(f\). NoisyNet, specifically, is a stochastic
process parameterized by its weights and their standard deviation. We will define
this parameterized function as \(g_\phi\), and their sampling distribution as
\(\rho_\phi \in \mathcal{P}\). Given a noise vector \(\xi\), we have
\(f(\bm{s}, \bm{a}) = g_\phi(\bm{a}, \bm{s}, \xi)\). This means that NoisyNet
gives a new policy function for each noise vector \(\xi\). Although this means
that NoisyNet can model stochastic policies, \citet{fortunato_noisy_2019} point
out that for the loss function it is optimising, there always exsists a deterministic
optimal policy. They show through emipirical analysis, however, that this does not
mean that the policy necessarily disregards the noise and converges to a deterministic
policy.

@@

\begin{multline}\label{eq:fkl}
    \nabla_\phi \KL[q_\phi(\f(\X)) \Vert p(\f(\X))] = \E_\rho \left[ \nabla_\phi \log \rho_\phi (\f(\X)) \right]\\
    + \E_\xi \left[ \nabla_\phi \f(\X)( \nabla_\f \log \rho_\phi(\f(\X)) - \nabla_\f \log p(\f(\X))) \right].
\end{multline}

@@

The first term is just the expected value of the score function, which is zero.
The difficult part in \eqref{eq:fkl} is to estimate \(\nabla_\f \log \rho(\f^\X)\)
and \(\nabla_\f \log p(\f^\X)\). \(\nabla_\f \log \rho(\f^\X)\) is likely 
intractable, considering \(\rho_\phi\) is a neural network with stochastic weights.
Depending on how we define the prior, however, \(\nabla_\f \log p(\f^\X)\)
can be easy to compute analytically. To reduce variance in the gradients, we
define tractable priors in this paper. To estimate the log-density gradient
\(\nabla_\f \log \rho(\f^\X)\), they use the Spectral Stein Gradient Estimator (SSGE)
\citep{shi_spectral_2018}. In short, this is a method for estimating the gradient
function of implicit distributions using approximations to eigenfunctions of a 
kernel-based operator. We bring three hyperparameters into the algorithm from this
method, the number of samples \(k\) from the implicit distribution used to approximate
the gradient, the number of eigenvectors \(J\) used to approximate the gradient, and
\(\eta\), a regularization parameter that smooths the gradient function.

@@

In order for this to match the functional ELBO and be a proper lower bound
for \(\log p(\D)\), \(\lambda\) should be set as \(\frac{1}{\lvert \D \rvert}\).
\cite{sun_functional_2019} note, however, that the minimization target is only a lower
bound for \(\KL \left[ \lvert \rho(f) \Vert p(f) \right]\), so a larger
value for \(\lambda\) is likely necessary to prevent overfitting. They use,
\(\lambda = \frac{1}{\lvert D_s \rvert}\), a value of one over the batch size instead.
This means that the minimization target becomes a proper lower bound of \(\log p(\D_s)\).

@@

\begin{equation}
    \mathcal{L} = \frac{1}{\lvert \D_s \rvert}\sum_{i=1}^{\lvert \D_s \rvert} \tau_i + \left( \mu_i - q_i \right)^2
    \exp (-\tau_i) + \lambda \KL[\rho \Vert p].
\end{equation}

@@

\begin{algorithm}
\caption{Functional Bayesian RL}
\begin{algorithmic}
    \Function{UpdatePosterior}{$\phi, \omega, \D$}
        \State \(T \sim \D\)
        \State \(\f \gets \{f_i \sim q_\phi \mid i=1,\dots,k\}\)
        \State \(\mathbf{b} \gets \{b_i \sim q_\omega \mid i=1,\dots,k\}\)
        \State \(\mathcal{F} \gets \emptyset\)
        \State \(\Delta_\mathcal{L} \gets 0\)
        \State \(S \gets 0\)
        \For{\(\{s,a,r,s'\} \in T\)}
            \State \(\bm{q_\mu}, \bm{q_\sigma} \gets \max_a \f(s,a)\)
            \State \(\mathcal{F} \gets \mathcal{F} \cup \f(s,a)\)
            \State \(\bm{G} \gets \{r + \gamma b_i(s,a) \mid i = 1, 
        \dots, k\}\)
            \State \(S \gets S + \frac{1}{k}p(\bm{q} \vert s, a)\)
            \State \(\Delta_\mathcal{L} \gets \Delta_\mathcal{L} - \frac{1}{\lvert k \rvert} \nabla_\phi \log p(\bm{q_\mu} \vert \bm{G}, \bm{q_\sigma})\)
        \EndFor
        \State \(\Delta_\KL \gets S - SSGE(\mathcal{F})\)
        \State \(\phi \gets \phi + \eta_\phi (\Delta_\mathcal{L} - \Delta_\KL)\)
        \State \(\omega \gets \omega + \eta_\omega (\omega - \phi)\)
    \EndFunction
\end{algorithmic}
\end{algorithm}

>>>>

Before we dwelve into the background, we need to presicise some notation. Most of the
theory will be based on stochastic functions. The stochastic process we are interested is
generating value functions on an MDP, and exists on the probability space on the probability
space \((\Omega, \mathcal{F}, P)\). It can be written as
\(\{Q(\bm{s}, \bm{a}) : (\bm{s}, \bm{a}) \in \state \times \action \}\). For
any \(\omega \in \Omega\), \(Q(\cdot, \cdot, \omega)\) is a sample function
mapping \(\state \times \action \rightarrow \mathbb{R}\). To simplify notation
we will denote sample functions as \(f\). For a set of multiple sampled functions we will
use \(\f\), where \(\f(\X)\) means each sampled function evaluated at the same point,
i.e. \(\{f(\X) \mid f \in \f\}\).

@@

\subsection{Noisy Networks}
NoisyNet, specifically, is a stochastic
process represented by a neural network with stochastic weights. We will define
this parameterized function as \(g_\phi\), and their sampling distribution as
\(\rho_\phi \in \mathcal{P}\). Given a noise vector \(\xi\), we have
\(f(\bm{s}, \bm{a}) = g_\phi(\bm{a}, \bm{s}, \xi)\). Although this means
that NoisyNet can model stochastic policies, \citet{fortunato_noisy_2019} point
out that for the loss function it is optimising, there always exsists a deterministic
optimal policy. They show through emipirical analysis, however, that this does not
mean that the policy necessarily disregards the noise and converges to a deterministic
policy.

@@

\begin{multline}\label{eq:fkl}
    \nabla_\phi \KL[q_\phi(\f(\X)) \Vert p(\f(\X))] = \E_\rho)\phi \left[ \nabla_\phi \log \rho_\phi (\f(\X)) \right]\\
    + \E_\xi \left[ \nabla_\phi \f(\X)( \nabla_\f \log \rho_\phi(\f(\X)) - \nabla_\f \log p(\f(\X))) \right].
\end{multline}

@@

The first term is just the expected value of the score function, which is zero.
The difficult part in \eqref{eq:fkl} is to estimate \(\nabla_\f \log \rho_\phi(\f(\X))\)
and \(\nabla_\f \log p(\f(\X))\). \(\nabla_\f \log \rho_\phi(\f(\X))\) is likely 
intractable, considering \(\rho_\phi\) is a neural network with stochastic weights.
Depending on how we define the prior, however, \(\nabla_\f \log p(\f(\X))\)
can be easy to compute analytically. To reduce variance in the gradients, we
define tractable priors in this paper. To estimate the log-density gradient
\(\nabla_\f \log \rho_\phi(\f(\X))\), they use the Spectral Stein Gradient Estimator (SSGE)
\citep{shi_spectral_2018}. In short, this is a method for estimating the gradient
function of implicit distributions using approximations to eigenfunctions of a 
kernel-based operator. We bring three hyperparameters into the algorithm from this
method, the number of samples \(k\) from the implicit distribution used to approximate
the gradient, the number of eigenvectors \(J\) used to approximate the gradient, and
\(\eta\), a regularization parameter that smooths the gradient function.

@@

In order for this to match the functional ELBO and be a proper lower bound
for \(\log p(\D)\), \(\lambda\) should be set as \(\frac{1}{\lvert \D \rvert}\).
\cite{sun_functional_2019} note, however, that the minimization target is only a lower
bound for \(\KL \left[ \lvert \rho_\phi(f) \Vert p(f) \right]\), so a larger
value for \(\lambda\) is likely necessary to prevent overfitting. They use,
\(\lambda = \frac{1}{\lvert D_s \rvert}\), a value of one over the batch size instead.
This means that the minimization target becomes a proper lower bound of \(\log p(\D_s)\).

@@

\begin{equation}
    \mathcal{L} = \frac{1}{\lvert \D_s \rvert}\sum_{i=1}^{\lvert \D_s \rvert} \tau_i + \left( \mu_i - q_i \right)^2
    \exp (-\tau_i) + \lambda \KL[\rho_\phi \Vert p].
\end{equation}

@@

\begin{algorithm}
\caption{Functional Bayesian RL}
\begin{algorithmic}
    \Function{UpdatePosterior}{$\phi, \omega, \D$}
        \State \(T \sim \D\)
        \State \(\f \gets \{f_i \sim q_\phi \mid i=1,\dots,k\}\)
        \State \(\mathbf{b} \gets \{b_i \sim q_\omega \mid i=1,\dots,k\}\)
        \State \(\mathcal{F} \gets \emptyset\)
        \State \(\Delta_\mathcal{L} \gets 0\)
        \State \(S \gets 0\)
        \For{\(\{s,a,r,s'\} \in T\)}
            \State \(\bm{q_\mu}, \bm{q_\sigma} \gets \max_a \f(s,a)\)
            \State \(\bm{q_\sigma^\prime} \gets \max (\bm{q_\sigma}, \sigma_t)\)
            \State \(\mathcal{F} \gets \mathcal{F} \cup \f(s,a)\)
            \State \(\bm{G} \gets \{r + \gamma b_i(s,a) \mid i = 1, 
        \dots, k\}\)
            \State \(S \gets S + \frac{1}{k}p(\bm{q} \vert s, a)\)
            \State \(\Delta_\mathcal{L} \gets \Delta_\mathcal{L} - \frac{1}{\lvert k \rvert} \nabla_\phi \log p(\bm{q_\mu} \vert \bm{G}, \bm{q_\sigma^\prime})\)
        \EndFor
        \State \(\Delta_\KL \gets S - SSGE(\mathcal{F})\)
        \State \(\phi \gets \phi + \eta_\phi (\Delta_\mathcal{L} - \Delta_\KL)\)
        \State \(\omega \gets \omega + \eta_\omega (\omega - \phi)\)
    \EndFunction
\end{algorithmic}
\end{algorithm}
